```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=9, fig.height=6) 
```

<!-- NOTE: make sure exercises in ../R-selfassess.md and answers here match :-) -->

# Exercise 0

## Load the data

```{r loading}
load(url("https://github.com/scgenomics/scgenomics.github.io/blob/main/docs/self-assess/self-assessment.rda?raw=true"))

ls() # Let's see what is there
```

The `load()` statement retrieves an
object called `meta.data`, which is a `data.frame` containing
cell-specific data. Rows are cells; columns are various characteristics.

## Exercise 1

How many cells and how many characteristics are there?

```{r ex1}
dim(meta.data)
## or
nrow(meta.data)
ncol(meta.data)
```

## Exercise 2

Column `celltypes` contains the inferred cell type of each cell. How
many cells have type 'Unknown'?

```{r ex2}
sum(meta.data$celltypes=="Unknown")
```

##  Exercise 3

What is the most commonly occurring type, and how many are there? (Hint:
use the `table()` function)

```{r ex3}
sort(decreasing=TRUE,table(meta.data$celltypes))[1]
```
##  Exercise 3

Create a plot number of transcripts (column `nUMI` versus numbers of
genes (column `nGene`) of only the cells in cluster `0` (column
`seurat_clusters`)

```{r ex4}
## using base graphics:
cluster <- 0 # this makes a bit more general
data <- meta.data[ meta.data$seurat_clusters==cluster, ]

## There are various other ways of course:
## data <- subset(meta.data, seurat_clusters==cluster)
##   or
## library(dplyr)
## data <- filter(meta.data, seurat_clusters==cluster)

plot(x=data$nGene, y=data$nUMI, log='xy', # using logarithmic axes here
     main=paste0('genes vs UMIs per cell in cluster ', cluster),
     xlab='number of genes',
     ylab='number of UMIs')

## using ggplot2
library(ggplot2)

ggplot(data = data, aes(x = nGene, y = nUMI)) +
  geom_point() +  scale_x_log10() +  scale_y_log10() + 
  labs(title = paste0('genes vs UMIs per cell in cluster ', cluster))

```
